---
title: "Report_1"
output: github_document
---

### Opening necessary in the project libraries

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(janitor)
library(lubridate)
library(tidyverse)
library(ggplot2)
library(slider)
library(patchwork)
library(readxl)

```

# Part 1 

+ Importing data

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Importing data

cheese_data <- read_csv("Resources/cheese_data.csv") %>%
  janitor::clean_names()

## Storing data for factory 1

cheese_data_f1 <- cheese_data %>%
  filter(factory %in% "f1")

## Storing data for factory 2

cheese_data_f2 <- cheese_data %>%
  filter(factory %in% "f2")

```

This report starts with importing the data from the csv file. Two sets of data are saved separately for Factory 1 and for Factory 2. 

# Part 2 

+ Cleaning the data
+ Managing missing values

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Transponing the table

cheese_data_long <- cheese_data %>%
  pivot_longer(
    cols = c("m1", "m2", "m3", "m4", "m5"),
    names_to = "measure",
    values_to = "result"
  )

## How many missing values have to be replaced?

miss_values <- cheese_data_long %>%
  filter(result == 0) %>%
  nrow()

## How many timepoints have to be discarded?

discard_timepoints_num <- cheese_data_long %>%
  group_by(factory, timepoint) %>%
  mutate(num_NA = sum(if_else(result == 0, 1, 0))) %>%
  filter(num_NA >= 2) %>%
  dplyr::select (timepoint) %>%
  distinct() %>%
  nrow()

## What timepoints have to be discarded?

discard_time <- cheese_data_long %>%
  group_by(factory, timepoint) %>%
  mutate(num_NA = sum(if_else(result == 0, 1, 0))) %>%
  filter(num_NA >= 2) %>%
  dplyr::select (timepoint) %>%
  distinct() %>%
  pull(timepoint)

## Discarding missing values

cheese_data_long <- cheese_data_long %>%
  filter(!(timepoint %in% discard_time))

## Replacing missing values

cheese_data_long <- cheese_data_long %>%
  group_by(factory, timepoint) %>%
  mutate(
    result = na_if(result, 0),
    AVG_result = mean(result, na.rm = TRUE),
    result = if_else(is.na(result), AVG_result, result)
  )
  
```

In the next step the data is transponed in order to carry out further calculations. Missing and uexpected values are detected.  
From the obtained results we can see that **`r miss_values`** missing values needed to be replaced. These values were replaced with average value calculated based on the same factory and timepoint.  
Additionally **`r discard_timepoints_num`** timepoints were discard.

# Part 3

+ Cleaning the data
+ Managing unexpected values

```{r echo=TRUE, message=FALSE, warning=FALSE}

## How many unexpected values are you able to detect?

unexpected_value <- cheese_data_long %>%
  filter(result > 65 | result < 15) %>%
  nrow()

## Checking in which factory unexcpected values appears

unexpected_value_factory <- cheese_data_long %>%
filter(result > 65 | result < 15) %>%
mutate(factory_num = str_sub (factory,2,2)) %>%
pull (factory_num)

## Replacing unexpected values with average for this group

cheese_data_long <-  cheese_data_long %>%
  mutate(
    result = if_else(result > 65 | result < 15, 0, result),
    result = na_if(result, 0),
    AVG_result = mean(result, na.rm = TRUE),
    result = if_else(is.na(result), AVG_result, result)
  ) 

```

Looking closer at unexpected values, so values lower than 15 and higher than 65. 
I was able to detect `r unexpected_value` unexpected value. This value appeared in the Factory `r unexpected_value_factory`. After identyfing unexcpected values, I recalculated average results and repleced them with the new average calculated based on the factory and timepoint.

# Part 4 

+ Computing monitoring statistics for factory f1

```{r echo=TRUE, message=FALSE, warning=FALSE}

# Storing the resulting values, xbar and Range, in a tibble named spc_data_f1.

spc_data_f1 <- cheese_data_long %>%
  filter(factory %in% "f1") %>%
  group_by(timepoint) %>%
  mutate(
    xbar = AVG_result,
    xmin = min(result),
    xmax = max(result),
    Range = (xmax - xmin)
  ) %>%
  ungroup() %>%
  dplyr::select (timepoint, xbar, Range,week, factory) %>%
  distinct()

## How many rows does the tibble spc_data_f1 have?

data_f1_rows <- spc_data_f1 %>%
  nrow()

## Computing the average of xbar and Range values.

avg_range_f1 <- spc_data_f1 %>%
  summarise(avg_range_f1 = round(mean(Range), 2)) %>%
  pull()

avg_xbar_f1 <- spc_data_f1 %>%
  summarise(avg_xbar_f1 = round(mean(xbar), 2)) %>%
  pull()

```

Going over to monitoring statistics for Factory 1:  
The spc_data_f1 tibble has **`r data_f1_rows`** rows.  
The average of xbar is **`r avg_xbar_f1`**.  
The average of range values is **`r avg_range_f1`**.

# Part 5 

+ Building Shewhart control charts for factory f1

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Importing data

xbar_constants <- read_csv("Resources/XbarR_constants.csv") %>%
  janitor::clean_names()

d3 <- xbar_constants %>%
  filter(n == 5) %>%
  pull(d3_2)

d4 <- xbar_constants %>%
  filter(n == 5) %>%
  pull(d4)

a2 <- xbar_constants %>%
  filter(n == 5) %>%
  pull(a2)

## Adding LCL and UCL values

LCL_range_f1 <- avg_range_f1 * d3
UCL_range_f1 <- avg_range_f1 * d4

LCL_xbar_f1 <- avg_xbar_f1 - a2 * avg_range_f1
UCL_xbar_f1 <- avg_xbar_f1 + a2 * avg_range_f1

## Producing an R chart (lineplot of Range vs. timepoint).

spc_data_f1 %>%
  ggplot(mapping = aes(x = timepoint, y = Range)) +
  geom_line(color = 'blue') +
  geom_point(color = 'blue') +
  geom_hline(yintercept = LCL_range_f1, color = "red") +
  geom_hline(yintercept = UCL_range_f1, color = "red") +
  geom_hline(yintercept = avg_range_f1, linetype = 'dashed') +
  theme_minimal()
```

In this part of the report Shewhart control charts are being presented.  
Starting with the graph above, we can see R chart that shows data for Factory 1. Red lines on the chart highlights Lower and Upper Control Limits and dashed line avarage range value for Factory 1. Each point on the graph presents range of values for specific timepoint. 

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Producing an xbar chart (lineplot of xbar vs. timepoint).

spc_data_f1 %>%
  ggplot(mapping = aes(x = timepoint, y = xbar)) +
  geom_line(color = 'blue') +
  geom_point(color = 'blue') +
  geom_hline(yintercept = LCL_xbar_f1, color = "red") +
  geom_hline(yintercept = UCL_xbar_f1, color = "red") +
  geom_hline(yintercept = avg_xbar_f1, linetype = 'dashed') +
  theme_minimal()

```

Now going over to Xbar chart we can see again data presented for Factory1. Red lines on the chart highlights Lower and Upper Control Limits and dashed line avarage value for Factory 1. Each point on the graph presents average value for specific timepoint.

# Part 6 

+ Highlighting special causes of variations

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Preparing function

all_gte_ref_value_f1 <- function(x)
  all(x >= avg_xbar_f1)
all_low_ref_value_f1 <- function(x)
  all(x <= avg_xbar_f1)

## Saving red points

red_point_f1 <- spc_data_f1 %>%
  mutate(
    tenlow = slide_lgl(
      .x = xbar,
      .f = all_low_ref_value_f1,
      .before = 9,
      .complete = TRUE
    ),
    tenhigh = slide_lgl(
      .x = xbar,
      .f = all_gte_ref_value_f1,
      .before = 9,
      .complete = TRUE
    )
  ) %>%
  filter(tenlow == "TRUE" |
           tenhigh == "TRUE" | xbar > UCL_xbar_f1 | xbar < LCL_xbar_f1)

## Creating control chart with highlights

xbar_f1 <- spc_data_f1 %>%
  ggplot(mapping = aes(x = timepoint, y = xbar)) +
  geom_line(color = 'blue') +
  geom_hline(yintercept = LCL_xbar_f1, color = "red") +
  geom_hline(yintercept = UCL_xbar_f1, color = "red") +
  geom_hline(yintercept = avg_xbar_f1, linetype = 'dashed') +
  geom_point() +
  geom_point(data = red_point_f1, color = "red") +
  theme_minimal() +
  labs(subtitle = "Process Chart Factory 1",
       x = "Measurement",
       y = "Result")

xbar_f1

```

In this part of the report special causes of variations have been highlighted. Specific point counts as a special cause if it's outside of the control limits or if there are ten or more consecutive points on one side of the centerline. Special causes are highlighted in red on the above chart. We can notice that for factory 1 we obtained 8 such cases.

# Part 7

+ Building Shewhart control charts for factory f2

```{r echo=TRUE, message=FALSE, warning=FALSE}

# Storing the resulting values, xbar and Range, in a tibble named spc_data_f2.

spc_data_f2 <- cheese_data_long %>%
  filter(factory %in% "f2") %>%
  group_by(timepoint) %>%
  mutate(
    xbar = AVG_result,
    xmin = min(result),
    xmax = max(result),
    Range = (xmax - xmin)
  ) %>%
  ungroup() %>%
  dplyr::select (timepoint, xbar, Range,week, factory) %>%
  distinct()

## Computing the average of xbar and Range values.

avg_range_f2 <- spc_data_f2 %>%
  summarise(avg_range_f2 = mean(Range)) %>%
  pull()

avg_xbar_f2 <- spc_data_f2 %>%
  summarise(avg_xbar_f2 = mean(xbar)) %>%
  pull()

## Adding LCL and UCL values

LCL_range_f2 <- avg_range_f2 * d3
UCL_range_f2 <- avg_range_f2 * d4

LCL_xbar_f2 <- avg_xbar_f2 - a2 * avg_range_f2
UCL_xbar_f2 <- avg_xbar_f2 + a2 * avg_range_f2

## Producing an R chart (lineplot of Range vs. timepoint).

spc_data_f2 %>%
  ggplot(mapping = aes(x = timepoint, y = Range)) +
  geom_line(color = 'blue') +
  geom_hline(yintercept = LCL_range_f2, color = "red") +
  geom_hline(yintercept = UCL_range_f2, color = "red") +
  geom_hline(yintercept = avg_range_f2, linetype = 'dashed') +
  theme_minimal()

```

Now we are moving to present Shewhart control charts for Factory 2.  
The graph above shows R chart. Red lines highlights Lower and Upper Control Limits and dashed line avarage range value for Factory 2. Each point on the graph presents range of values for specific timepoint. 

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Preparing function

all_gte_ref_value_f2 <- function(x)
  all(x >= avg_xbar_f2)
all_low_ref_value_f2 <- function(x)
  all(x <= avg_xbar_f2)

## Saving red points

red_point_f2 <- spc_data_f2 %>%
  mutate(
    tenlow = slide_lgl(
      .x = xbar,
      .f = all_low_ref_value_f2,
      .before = 9,
      .complete = TRUE
    ),
    tenhigh = slide_lgl(
      .x = xbar,
      .f = all_gte_ref_value_f2,
      .before = 9,
      .complete = TRUE
    )
  ) %>%
  filter(tenlow == "TRUE" |
           tenhigh == "TRUE" | xbar > UCL_xbar_f2 | xbar < LCL_xbar_f2)

## Producing an xbar chart (lineplot of xbar vs. timepoint).

xbar_f2 <- spc_data_f2 %>%
  ggplot(mapping = aes(x = timepoint, y = xbar)) +
  geom_line() +
  geom_hline(yintercept = LCL_xbar_f2, color = "red") +
  geom_hline(yintercept = UCL_xbar_f2, color = "red") +
  geom_hline(yintercept = avg_xbar_f2, linetype = 'dashed') +
  geom_point() +
  geom_point(data = red_point_f2, color = "red") +
  theme_minimal() +
  labs(subtitle = "Process Chart Factory 2",
       x = "Measurement",
       y = "Result")

xbar_f2
```

On the Xbar chart above we can see again data for Factory 2. Red lines on the chart highlights Lower and Upper Control Limits and dashed line avarage value for Factory 1. Each point on the graph presents average value for specific timepoint. We can notice right away here that we received many more special causes than in case of Factory 1.

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Do we observe similar patterns in xbar charts of both factories?

xbar_f1 / xbar_f2

```

Now it's time to compare results between two these two factories.  
We can see that patterns for this two factories are different. Factory 1 has results more spreaded per one timepoint, while for Factory 2 values are more cummulated. In general values for Factory 1 devates more than for Factory 2, which could mean that process for Factory 2 is more stable.

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Producing a scatterplot of xbar values vs. week for both factories f1 and f2.

scatter_data <- cheese_data_long %>%
  group_by(timepoint, week, factory) %>%
  mutate(
    xbar = AVG_result,
     factory = if_else(factory == "f1", "Factory 1", "Factory 2"),
    week = as.factor(week)
  ) %>%
  ungroup() %>%
  dplyr::select (timepoint, xbar, week, factory) %>%
  distinct()

ggplot() +
  geom_point(scatter_data, mapping = (aes(
    x = week, y = xbar, color = factory
  ))) +
  theme_minimal() +
  geom_point(data = red_point_f1 %>%
                mutate(factory = if_else(factory == "f1", "Factory 1", "Factory 2")),
             mapping = aes(x = week, y = xbar), color = "black") +
  geom_point(data = red_point_f2 %>%
              mutate(factory = if_else(factory == "f1", "Factory 1", "Factory 2")),
             mapping = aes(x = week, y = xbar), color = "black") +
  facet_grid( vars(row = factory)) +
  labs(title = "Scatterplot of xbar values vs. week",
       x = "Week",
       y = "Result") +
  theme(legend.position = "none")
  
  
```

From the obtained scatterplot we can confirm previous assumptions that values for Factory 2 are cummulated and for Factory 1 more spread.  
We can see as well that for Factory 2 we have many more special causes.

# Part 8 

+ Redefining monitoring data for factory f2

```{r echo=TRUE, message=FALSE, warning=FALSE}

week_avg_f2 <- cheese_data_long %>%
  ungroup() %>%
  filter(factory %in% "f2") %>%
  group_by(week) %>%
  summarise(avg_week = mean(result)) %>%
  dplyr::select(week, avg_week)

M <- week_avg_f2 %>%
  summarise(M = mean(avg_week)) %>%
  pull(M)

MR_bar <- week_avg_f2 %>%
  mutate(prev_avg = lag(avg_week),
         MR = abs(avg_week - prev_avg)) %>%
  summarise(MR_bar = mean(MR, na.rm = TRUE)) %>%
  pull(MR_bar)

LCL_mr <- M - 2.66 * MR_bar
UCL_mr <- M + 2.66 * MR_bar


week_avg_f2  %>%
  ggplot(mapping = aes(x = week, y = avg_week)) +
  geom_line(color = 'blue') +
  geom_hline(yintercept = LCL_mr, color = "red") +
  geom_hline(yintercept = UCL_mr, color = "red") +
  geom_hline(yintercept = M , linetype = 'dashed') +
  geom_point(color = 'blue') +
  theme_minimal() +
  labs(title = "Weekly Process Chart Factory 2",
       x = "Measurement",
       y = "Week")

```

Above chart presents redifined weekly average values for Factory 2. Here again red lines highlights Lower and Upper Control Limits and dashed line avarage value.

# Part 9 - open ended question

+ Producing a control chart using my own data

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Importing data

cow_data <- read_xlsx("Resources/milk_cow.xlsx") %>%
  janitor::clean_names()

## Clearing dates

cow_data <- cow_data %>%
  mutate(date = date(date_time),
         time = hour(date_time)) %>%
  dplyr:: select(date,
                 milk_prod = "milk_prod_per_cow_kg",
                 time)

# Storing the resulting values, xbar and Range

spc_cow_data <- cow_data %>%
  group_by(date) %>%
  mutate(
    xbar = mean(milk_prod, na.rm = TRUE),
    xmin = min(milk_prod),
    xmax = max(milk_prod),
    Range = (xmax - xmin)
  ) %>%
  ungroup() %>%
  dplyr::select (date, xbar, Range) %>%
  distinct()


## Computing the average of xbar and Range values.

avg_range_cow <- spc_cow_data %>%
  summarise(avg_range_cow = mean(Range)) %>%
  pull()

avg_xbar_cow <- spc_cow_data %>%
  summarise(avg_xbar_cow = mean(xbar)) %>%
  pull()

## Importing xbar constants

b3 <- xbar_constants %>%
  filter(n == 2) %>%
  pull(d3_2)

b4 <- xbar_constants %>%
  filter(n == 2) %>%
  pull(d4)

g2 <- xbar_constants %>%
  filter(n == 2) %>%
  pull(a2)

## Adding LCL and UCL values

LCL_range_cow <- avg_range_cow * b3
UCL_range_cow <- avg_range_cow * b4

LCL_xbar_cow <- avg_xbar_cow - g2 * avg_range_cow
UCL_xbar_cow <- avg_xbar_cow + g2 * avg_range_cow

## Preparing inline text

min_date <- cow_data %>%
  summarise(min(date))

max_date <- cow_data %>%
  summarise(max(date))

min_time <- cow_data %>%
  summarise(min(time))

max_time <- cow_data %>%
  summarise(max(time))

```

Presented in this part data comes from this [GitHub Account](https://github.com/ourcodingclub/CC-time-series).  
In this data set we have information about cow milking done in the traditional way.  
This measurement were taken between `r min_date` and `r max_date`.
For every day of this time period we have two results, from `r min_time` and `r max_time`.

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Producing an R chart 

spc_cow_data %>%
  ggplot(mapping = aes(x = date, y = Range)) +
  geom_line(color = 'blue') +
  geom_point(color = 'blue') +
  geom_hline(yintercept = LCL_range_cow, color = "red") +
  geom_hline(yintercept = UCL_range_cow, color = "red") +
  geom_hline(yintercept = avg_range_cow, linetype = 'dashed') +
  theme_minimal()
```

From the obtained chart we can see that values are crossing slightly the LCL range. As the traditional milking process is not that stable as the one in the factory we can notice more spread values than on the previous charts for Factory 1 and 2. 

```{r echo=TRUE, message=FALSE, warning=FALSE}

## Producing an xbar chart 

spc_cow_data %>%
  ggplot(mapping = aes(x = date, y = xbar)) +
  geom_line(color = 'blue') +
  geom_point(color = 'blue') +
  geom_hline(yintercept = LCL_xbar_cow, color = "red") +
  geom_hline(yintercept = UCL_xbar_cow, color = "red") +
  geom_hline(yintercept = avg_xbar_cow, linetype = 'dashed') +
  theme_minimal()
```

From the obtained chart we can noticed that values are changing over the time but are almost always within the range. Here again compering to results from factory it seems like this traditional process is less stable.
